---
title: "Supplemental Figures: Paired compositional similarity vs. Environment"
geometry: margin=2.54cm
date: "November 7, 2015"
output: pdf_document
header-includes: \usepackage{array}
---


## Species spatial abundance distribution (SSAD)

The SSAD represents the distribution of a species' abundance among sites. In 
contrast, the species abundance distribution (SAD) is the distribution of a 
site's abundance among species. Here, we compare each OTU's two SSADs (one from 'all' and one from 'active').

\pagebreak

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
# Retrieve and set the working directory
rm(list=ls())
getwd()
setwd("~/GitHub/Dimensions/Aim3/papers/Paired")
```


```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
# Load packages
require("picante")
require("ape")
require("vegan") # biodiversity estimators and related functions
require("fossil")
require("simba")
require("reshape")
require(reshape)
require(IDPmisc)
source("~/GitHub/Dimensions/Aim3/bin/MothurTools.R")
source("~/GitHub/Dimensions/Aim3/bin/metrics.R")
load(file = "~/GitHub/Dimensions/Aim3/Mothur/INPond_Initial.RData")
```


```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
# Load Environmental and Geographical Data
env <- read.table("~/GitHub/Dimensions/Aim3/DATA/EnvData/20130801_PondDataMod.csv",
                  sep = ",", header = TRUE)

env <- subset(env, chla < 2000)
env <- subset(env, pH > 1)
env <- subset(env, Salinity > 0.0)
env <- subset(env, TDS < 5.0)

env.dat <- as.matrix(subset(env, select = Diameter:DON))

# Select DNA Data: Use the `grep()` Command and Rename with `gsub()`
# The active portion, based on cDNA 
active.comm <- Pond97[grep("*-cDNA", rownames(Pond97)), ]
rownames(active.comm) <- gsub("\\-cDNA", "", rownames(active.comm))
rownames(active.comm) <- gsub("\\_", "", rownames(active.comm))

# The community without respect to active or not, 16S rRNA gene sequences
all.comm <- Pond97[grep("*-DNA", rownames(Pond97)), ]
rownames(all.comm) <- gsub("\\-DNA", "", rownames(all.comm))
rownames(all.comm) <- gsub("\\_", "", rownames(all.comm))

# Remove Sites Not in the Environmental Data Set
active.comm <- active.comm[rownames(active.comm)  %in% env$Sample_ID, ]
all.comm <- all.comm[rownames(all.comm)  %in% env$Sample_ID, ]

# Remove Zero-Occurrence Taxa 
#active.comm <- active.comm[ , colSums(active.comm) > 0]
#all.comm <- all.comm[ , colSums(all.comm) > 0]
```


```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
# Get Dominant members
cutoff <- 100

all.sub.comm <- all.comm[, 1:cutoff]
active.sub.comm <- active.comm[, 1:cutoff]
#all.sub.comm <- all.comm
#active.sub.comm <- active.comm
```


Paired-SSADs of the 4 most dominant OTUs are pretty similar. Does this change as
OTUs become more rare?

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
par(mfrow=c(2, 2))

ct <- 1         # a counter variable
while (ct <= 4){ # the while statement
  ad.all <- all.comm[,ct] # choose an OTU
  ad.all <- ad.all[ad.all > 0]
  kdens.all <- density(log(ad.all))
  max.all <- max(kdens.all$y)
  
  ad.active <- active.comm[,ct]
  ad.active <- ad.active[ad.active > 0]
  kdens.active <- density(log(ad.active))
  max.active <- max(kdens.active$y)
  
  max.y <- max(c(max.all, max.active))
  c <- 'red' #sample(colours(), 1)
  
  plot(kdens.all, col = c, xlab='log(site abundance)', ylim=c(0, max.y),
  ylab='Probability density', main=paste('SSAD for otu ',ct))
  lines(kdens.active, col = c, lty=2)
  ct <- ct + 1
  }
```

\pagebreak
Often, the rare-biosphere begins to rear its little head by the 10th OTU or so. But, 
here we see that paired-SSADs of the 11 to 14th most dominant OTUs are pretty similar. 

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
par(mfrow=c(2, 2))

ct <- 11       # a counter variable
while (ct <= 14){ # the while statement
  ad.all <- all.comm[,ct] # choose an OTU
  ad.all <- ad.all[ad.all > 0]
  kdens.all <- density(log(ad.all))
  max.all <- max(kdens.all$y)
  
  ad.active <- active.comm[,ct]
  ad.active <- ad.active[ad.active > 0]
  kdens.active <- density(log(ad.active))
  max.active <- max(kdens.active$y)
  
  max.y <- max(c(max.all, max.active))
  c <- 'blue' #sample(colours(), 1)
  
  plot(kdens.all, col = c, xlab='log(site abundance)', ylim=c(0, max.y),
  ylab='Probability density', main=paste('SSAD for otu ',ct))
  lines(kdens.active, col = c, lty=2)
  ct <- ct + 1
  }
```

\pagebreak
Likewise, paired-SSADs of the 101 to 104th most dominant OTUs are pretty similar. 

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
par(mfrow=c(2, 2))

ct <- 101       # a counter variable
while (ct <= 104){ # the while statement
  ad.all <- all.comm[,ct] # choose an OTU
  ad.all <- ad.all[ad.all > 0]
  kdens.all <- density(log(ad.all))
  max.all <- max(kdens.all$y)
  
  ad.active <- active.comm[,ct]
  ad.active <- ad.active[ad.active > 0]
  kdens.active <- density(log(ad.active))
  max.active <- max(kdens.active$y)
  
  max.y <- max(c(max.all, max.active))
  c <- 'magenta' #sample(colours(), 1)
  
  plot(kdens.all, col = c, xlab='log(site abundance)', ylim=c(0, max.y),
  ylab='Probability density', main=paste('SSAD for otu ',ct))
  lines(kdens.active, col = c, lty=2)
  ct <- ct + 1
  }
```

\pagebreak
And so are paired-SSADs of the 1001 to 1004th most dominant OTUs! 

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
par(mfrow=c(2, 2))

ct <- 1001       # a counter variable
while (ct <= 1004){ # the while statement
  ad.all <- all.comm[,ct] # choose an OTU
  ad.all <- ad.all[ad.all > 0]
  kdens.all <- density(log(ad.all))
  max.all <- max(kdens.all$y)
  
  ad.active <- active.comm[,ct]
  ad.active <- ad.active[ad.active > 0]
  kdens.active <- density(log(ad.active))
  max.active <- max(kdens.active$y)
  
  max.y <- max(c(max.all, max.active))
  c <- 'darkgreen' #sample(colours(), 1)
  
  plot(kdens.all, col = c, xlab='log(site abundance)', ylim=c(0, max.y),
  ylab='Probability density', main=paste('SSAD for otu ',ct))
  lines(kdens.active, col = c, lty=2)
  ct <- ct + 1
  }
```

\pagebreak
Let's examine whether percent differences in the paired-SSADs change as a 
function of rank in the site-by-species matrix.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=6}
par(mfrow=c(2, 2))

ranks <- c()
ks.vals <- c()
medians <- c()
means <- c()
modes <- c()
skews <- c()
vars <- c()

ct <- 1       # a counter variable
while (ct <= 180){ # the while statement
  ad.all <- all.comm[,ct] # choose an OTU
  ad.all <- ad.all[ad.all > 0]
  
  ad.active <- active.comm[,ct]
  ad.active <- ad.active[ad.active > 0]
  
  diff <- abs(var(ad.all) - var(ad.active))
  #diff <- var(ad.all) - var(ad.active)
  avg <- mean(c(var(ad.all), var(ad.active)))
  vars <- c(vars, (diff/avg)*100)
  
  diff <- abs(median(ad.all) - median(ad.active))
  #diff <- median(ad.all) - median(ad.active)
  avg <- mean(c(median(ad.all), median(ad.active)))
  medians <- c(medians, (diff/avg)*100)
  
  diff <- abs(mean(ad.all) - mean(ad.active))
  #diff <- mean(ad.all) - mean(ad.active)
  avg <- mean(c(mean(ad.all), mean(ad.active)))
  means <- c(means, (diff/avg)*100)
  
  d.all = density(ad.all)
  d.active = density(ad.active)
  i.all = which.max(d.all$y)
  i.active = which.max(d.active$y)
  diff <- abs(i.all - i.active)
  #diff <- i.all - i.active
  avg <- mean(c(i.all, i.active))
  modes <- c(modes, (diff/avg)*100)
  
  ranks <- c(ranks, ct)
  ct <- ct + 1
  }

plot(ranks, medians, col = 'red', xlab='Rank in Site-by-Species Matrix',
     ylab='% Difference', main='Median site abundance')

plot(ranks, means, col = 'blue', xlab='Rank in Site-by-Species Matrix',
     ylab='% Difference', main='Mean site abundance')

plot(ranks, modes, col = 'magenta', xlab='Rank in Site-by-Species Matrix',
     ylab='% Difference', main='Modal site abundance')

plot(ranks, vars, col = 'green', xlab='Rank in Site-by-Species Matrix', 
     ylab='% Difference', main='Variance in site abundance')

```