---
title: "Flow cytometry analysis of active, dormant, and dead microbial populations in Indiana Ponds (South Central Indiana)"
author: "M.L. Larsen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
fontsize: 11pt
geometry: margin = 0.75in
---

# Project Overview

The microbes within communities have wide ranging metabolic activities that are influenced by resource availability.
We can classify their metabolic activity with DNA and RNA content. Here, we seek to enumerate the fractions of the active (greater RNA than DNA), dormant (less RNA to DNA), and dead populations of microbes within freshwater ecosystems.

##Data Collection

Data for this project were collected as part of a long term sampling regime on University Lake in Bloominton, Indiana and wildlife refuge ponds scattered through south central Indiana. All data were collected in the Indiana University Flow Cytometry Core Facility operated by Christiane Hassel and processed by M.L. Larsen.

##Document workflow
1. Setup work environment

___

## WORK ENVIRONMENT
```{r work-environment, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
# Setup working environment
rm(list = ls())
setwd("C:/Users/Megan/Github/Dimensions/Aim3")

# Load packages

# Load data
env.dat <- read.csv("", header = TRUE)
flow.dat <- read.csv("", header = TRUE)

# Output directories
data.output <- "./DATA/FlowCytometryData/"
```

## Indiana Ponds Geographic Abundances

### Population relative abundances
```{r INPonds, include=FALSE, eval=FALSE}


```

\clearpage

## R Environment and package information

All analyses were completed with `r sessionInfo()$R.version$version.string` and the following packages: 

```{r session-info}
sessioninfo()
```

\clearpage

## References