---
title: "Brown County Ponds"
author: "Ken Locey"
date: "January 27, 2015"
output: pdf_document
---

The purpose of this document is to **explore** patterns and relationships of biodiversity and biogeography for a set of 52 pond samples from central Indiana, mainly Brown Country. Empty R chunks signify areas to be filled.

## 1.) SETUP
### Retrieve and Set Your Working Directory

```{r, results = 'hide'}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/Dimensions/Aim3") 
```

### Basic packages

```{r}
#install.packages("sp")  # classes for spatial data
#install.packages("raster")  # grids, rasters
#install.packages("rasterVis")  # raster visualisation
#install.packages("maptools")
#install.packages("rgeos", type="source") # and their dependencies# and their dependencies
#install.packages("RgoogleMaps")
#install.packages("googleVis")
#install.packages("BiodiversityR", type = "source", dependencies=TRUE)

require("BiodiversityR")
require("vegan")

require("sp")  # classes for spatial data
require("raster")  # grids, rasters
require("rasterVis")  # raster visualisation
require("maptools")
require("rgeos") # and their dependencies
require("dismo")
require("RgoogleMaps") # overlays of data onto a Google map.
require("googleVis")
```

## 2.) Mapping

### Map sample ponds onto Google map
Retrieve base maps from Google with the `gmap` function in Species distribution modeling package `dismo`.

```{r}
# Map centered on Brown County
newmap <- GetMap(center = c(39.1,-86.3), zoom = 8, destfile = "BrownComap.png", maptype="terrain")

PlotOnStaticMap(newmap, zoom = 8, cex = 2, col='blue')
Ponds <- read.table(file="BrownCoData/20130801_PondData.csv", head=TRUE, sep=",")

lats <- Ponds[, 3]
lons <- Ponds[, 4]
PlotOnStaticMap(newmap, lats, lons, zoom=8, cex=1.5, col='red', add = FALSE)
```


### Plot sample ponds on web-page to search, scroll, magnify

```{r}
names <- Ponds[, 1]
latlongs <- transform(names, latlongs = interaction(lats, lons, sep=':'))

M1 <- gvisMap(latlongs, "latlongs", "X_data", options = list(showTip=TRUE, showLine=TRUE, enableScrollWheel=TRUE, mapType='satellite', useMapTypeControl=TRUE, width=800,height=600))
plot(M1)
```


## 3.) Simple relationships of environment to demographics

### Environmental parameters

```{r}
EnvData <- Ponds[5:21]
Temp <- EnvData[6]
```


### A function for calculating OTU richness

```{r}
OTU.obs <- function(x = ""){
  rowSums(x > 0) * 1
  }
```


### Generate data object for S and N of dormant and active across samples

```{r}
# Active
Active <- read.csv(file="BrownCoData/active.csv", head=TRUE, sep=",")
Active <- Active[-1] # remove first column (names)
Active.S <- as.matrix(OTU.obs(Active))
Active.N <- as.matrix(rowSums(Active)) # no. reads

# Dormant
Dormant <- read.csv(file="BrownCoData/dormant.csv", head=TRUE, sep=",")
Dormant <- Dormant[-1] # remove first column (names)
Dormant.S <- as.matrix(OTU.obs(Dormant))
Dormant.N <- as.matrix(rowSums(Dormant)) # no. reads
```


### Scatter plots and correlation of sample-based N and S
Because regardless of what they mean, Ken is always thinking about N and S.

```{r}
A.dem <- cbind(Active.S, Active.N)
labels <- c('richness', 'abundance')
pairs(A.dem, labels)
cor <- cor(A.dem)
cor

mydata <- as.data.frame(Active.N)
fit <- lm(Active.S ~ Active.N, data=mydata)
summary(fit)

Ddem <- cbind(Dormant.S, Dormant.N)
labels <- c('richness', 'abundance')
pairs(Ddem, labels)
cor <- cor(Ddem)
cor

mydata <- as.data.frame(Dormant.N)
fit <- lm(Dormant.S ~ Dormant.N, data=mydata)
summary(fit)

AvD.S <- cbind(Active.S, Dormant.S)
labels <- c('Active.S', 'Dormant.S')
pairs(AvD.S, labels)
cor <- cor(AvD.S)
cor


AvD.N <- cbind(Active.N, Dormant.N)
labels <- c('Active.N', 'Dormant.N')
pairs(AvD.N, labels)
cor <- cor(AvD.N)
cor
```


### Simple sample demographics vs. Environmental variables 

```{r}
# Dormant richness vs. Temp
SvTemp <- cbind(Dormant.S, Temp)
labels <- c('Dormant.S', 'Temp')
pairs(SvTemp, labels)
cor <- cor(SvTemp)
cor

mydata <- as.data.frame(Temp)
fit <- lm(Dormant.S ~ Temp, data=mydata)
summary(fit) # kinda sorta

# Active richness vs. Temp 
SvTemp <- cbind(Active.S, Temp)
labels <- c('Active.S', 'Temp')
pairs(SvTemp, labels)
cor <- cor(SvTemp)
cor

mydata <- as.data.frame(Temp)
fit <- lm(Active.S ~ Temp, data=mydata)
summary(fit) # wonk wonk

```


### Messing around with Temperature, Salinity, and Temporo-mandibular Syndrome (TDS)  

```{r}
fit <- lm(Dormant.S ~ Temp, data= EnvData)
summary(fit)

fit <- lm(Active.S ~ Salinity + TDS, data= EnvData)
summary(fit)
```


## 4.) alpha-Diversity

### Evenness

```{r}

```


### Diversity

```{r}

```


### Local richness Estimation  
A function using vegan's `estimateR` function

```{r}
S.est <- function(x = ""){
  est <- estimateR(x)
  chao1 <- est[2]
  ace <- est[4]
  return(chao1)
}
```


Plotting kernel density curves for estimated Richness

```{r}
par(mfrow=c(1, 2))

Active.S.est <- apply(Active, 1, S.est)
Dormant.S.est <- apply(Dormant, 1, S.est)

ASx <- density(Active.S.est)
DSx <- density(Dormant.S.est)

plot(ASx, main = "Est. Richness Active, Chao1", col = 'DarkCyan')
plot(DSx, main = "Est. Richness Dormant, Chao1", col = 'magenta')
```


### Coverage

A functions for coverage (C)

```{r}
# Good's Coverage == Bad Logic
Goods <- function(x = ""){
  1.0 - (sum(x == 1) / sum(x))
  }
```


Plotting kernel density curves for coverage

```{r}
par(mfrow=c(1, 2))

Active.C <- apply(Active, 1, Goods)
summary(Active.C)
Ax <- density(Active.C)

Dormant.C <- apply(Dormant, 1, Goods)
summary(Dormant.C)
Dx <- density(Dormant.C)

plot(Ax, main = "Good's C for Active", col = 'DarkCyan')
plot(Dx, main = "Good's C for Dormant", col = 'magenta')
```


### Rarefaction

Plotting rarefaction curves for each site's Active and Dormant

```{r}
par(mfrow=c(1,2))
min.N <- min(rowSums(Active))    
S.rarefy <- rarefy(x = Active, sample = min.N, se = TRUE)   
rarecurve(x = Active, main = "Active Richness", step = 5000, col = "DarkCyan", cex = 0.6, las=1)

min.N <- min(rowSums(Dormant))    
S.rarefy <- rarefy(x = Dormant, sample = min.N, se = TRUE)   
rarecurve(x = Dormant, main = "Dormant Richness", step = 1000, col = "magenta", cex = 0.6, las=1)
```


## 5.) beta-Diversity

```{r}
# INSERT NIFTY ANALYSES OF UNIVARIATE, DISTANCE, AND DISSIMILARITY METRICS 
```


## 6.) Gamma Gamma Gamma-Diversity

### Richness Estimation (gamma)

```{r}
Active.gamma <- specpool(Active)
Active.gamma

Dormant.gamma <- specpool(Dormant)
Dormant.gamma
```


### OTU-accumulation curves (gamma)

For Active
```{r}
x <- specaccum(Active, method = "random", permutations = 20, ci.type = "polygon", conditioned = TRUE, gamma = "jack1")

plot(x, add = FALSE, ci = 2, ci.type = "polygon", col = "DarkCyan", ci.col = col, ci.lty = 1, xlab = "Sites", ylab = x$method, main = "Active gamma-diversity")

boxplot(x, add = FALSE)
```


For Dormant
```{r}
x <- specaccum(Dormant, method = "random", permutations = 20, ci.type = "polygon", conditioned = TRUE, gamma = "jack1")

plot(x, add = FALSE, ci = 2, ci.type = "polygon", col = "magenta", ci.col = col, ci.lty = 1, xlab = "Sites", ylab = x$method, main = "Dormant gamma-diversity")

boxplot(x, add = FALSE)
```


## 7.) Macroecological Patterns and Relationships

### Species-abundance distributions

Using function in `vegan`
```{r}
Active.RACresults <- radfit(Active)
#Active.RACresults
plot(Active.RACresults)

Dormant.RACresults <- radfit(Dormant)
#Dormant.RACresults
plot(Dormant.RACresults, las=1, cex.lab = 1.4, cex.axis = 1.25)
```


### Species-area relationships

```{r}


```


### Spatial-abundance distributions

```{r}

```


### Rarity, Dominance, and Evenness vs. N

```{r}

```

