---
title: "Long-Term Dormancy"
author: "Stuart E. Jones and Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

## OVERVIEW
In January 2013, we started an experiment using ~24 strains of bacteria -- some "lab" strains others enviornmental strains -- to estimate survivorship during long-term starvation. Each strain was grown un into logarithmic phase and harvested. The cells were pelleted and washed 5x before being put in replicate 50 mL Falcon tubes. Subsamples were taken on overtime and plated onto R2A plates for enumeration as CFUs. After looking at plots of the data, in became apparent that CFUs were declining at a non-constant rate on semi-log plots. Here, we test whether or not the Weibull function can help explain variation in death curves.  

## 1) SET WORKING DIRECTORY AND LOAD DATA
```{r}
## Retrieve and Set Your Working Directory
rm(list = ls())
getwd()
setwd("~/GitHub/Dimensions/Aim1/DeathCurves/")
```

##2) LOAD DATA
Using KBS0703 as a test. 
Ran into problems where following error was generated:
"initial value in 'vmmin' is not finite"
I think this was due to having time points where time = 0

```{r, results='hide', warning=FALSE, message=FALSE}
KBS0703 <- read.table("data/KBS0703.txt", header = TRUE)
KBS0703 <- data.frame(KBS0703[,4],KBS0703[,9:10])
logabund <- log10(KBS0703$Abund)
KBS0703 <- data.frame(KBS0703, logabund)
KBS0703$time[KBS0703$time == 0] <- 0.1 # replace time 0 with time 0.1
#KBS0703 <- KBS0703[ which(KBS0703$time > 0), ] # removes time = 0
colnames(KBS0703) <-c("rep","abund","time","logabund")
```

## 3) WEIBULL FUNCTION
```{r, results='hide', warning=FALSE, message=FALSE}

# Initial parameters
A = 200 # Initial death (larger = slower) 
B = 1 # Bend (upper = 1 = first-order decay)
C = round(max(KBS0703$logabund),1) # intercept
Z = 10 # Error

# Function
fit <- mle2(logabund ~ dnorm(mean = c * (time / a)^(b-1) * exp(-1*(time/a)^b), sd = z), 
            start = list(a = A, b = B, c = C, z = Z), data = KBS0703) 
```

## 4)  PLOT DATA AND FUNCTION
```{r}
par(mar = c(5, 7, 5, 7))

plot(KBS0703$time, KBS0703$logabund, pch = 22, 
          bg = "white", col = "black", lwd = 2, cex.axis = 1.25, xlab = NA, ylab = NA,
          xlim = c(-50, 1000), ylim = c(3, 9), yaxt = "n", xaxt = "n")
box(lwd = 2)

# Major Axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("10^3", "10^4", "10^5", "10^6", "10^7", "10^8", "10^9"), at = c(3, 4, 5, 6, 7, 8, 9))
 
axis(side = 4, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    at=c(3, 4, 5, 6, 7, 8, 9), labels = F)
 
axis(side = 1, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c(0, 250, 500, 750, 1000), at = c(0, 250, 500, 750, 1000))

axis(side = 3, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    at = c(0, 250, 500, 750, 1000), labels = F)

# Plot function
curve(coef(fit)[[3]] * (x / coef(fit)[[1]])^(coef(fit)[[2]]-1) * exp(-1*(x/coef(fit)[[1]])^coef(fit)[[2]]), 
      from = 0.1, to = 1000, add = TRUE, lty = 2, lwd = 4, col = "red") 
```