---
title: "Sample Map"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---


## 1.) SETUP
### A. Retrieve and Set Your Working Directory

```{r, results = 'hide'}
rm(list=ls())
getwd()
setwd("~/GitHub/Dimensions/Aim3")
```

### B. Load Packages 

We will use the `vegan` package for biodiversity estimators and related functions.

```{r, results = 'hide', message = FALSE, warning = FALSE} 
require("vegan")
```

Be sure to run **install.packages(PackageName)**, if not previously installed, where PackageName is the name of the package you want installed.
Or, run **install.packages(PackageName, type="source", dependencies=TRUE)**, if the previous command doesn't work.

```{r, results = 'hide', message = FALSE, warning = FALSE} 
require("sp")          # Classes and methods for handling spatial data
require("geoR")        # Methods for geostatistical analyses
require("rgdal")       # Geospatial Data Abstraction Library
require("raster")      # Methods to create a RasterLayer object
require("RgoogleMaps") # For querying the Google server for static maps. 
require("maptools")    # Tools for manipulating and reading geospatial data
```

### C. Load and Compile a Large Dataset

```{r, results = 'hide', message = FALSE, warning = FALSE} 
Ponds <- read.table("~/GitHub/Dimensions/Aim3/DATA/EnvData/20130801_PondDataMod.csv", sep = ",", header = TRUE)

Ponds <- subset(Ponds, pH > 1)
Ponds <- subset(Ponds, long < -86.35)
Ponds <- subset(Ponds, lat > 39.05)
Ponds <- subset(Ponds, TDS < 5.0)

lats <- as.numeric(Ponds[, 3]) # latitudes (north and south)
lons <- as.numeric(Ponds[, 4]) # longitudes (east and west)
```


## 2.) MAP SAMPLES AND DATA
Let's visualize the spatial distribution of our samples with a basic map in RStudio.
Let's generate a map of the refuge ponds using the `GetMap` function in the package `RgoogleMaps`.
This map will be centered on Brown County, Indiana (39.1 latitude, -86.3 longitude).

```{r}
newmap <- GetMap(center = c(39.1,-86.3), zoom = 10, destfile = "PondsMap.png",
                 maptype="terrain")
PlotOnStaticMap(newmap, zoom = 10, cex = 2, col='blue') # Plot map in RStudio
PlotOnStaticMap(newmap, lats, lons, cex=1, pch=20, col='red', add = TRUE)
```
